---
title: "Homework Template"
output:
  html_document: default
  pdf_document: default
date: "2023-10-09"
---

<style type="text/css">
  body{
  font-size: 12pt;
  font-family: "Arial";
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
# Installs and loads the required packages for this project

# Vector of the packages used
packages <- c("palmerpenguins", "tidyverse")

# Installs only the packages which are not already downloaded
install.packages(setdiff(packages, rownames(installed.packages())), dependencies = TRUE)
# Library all of the packages required
lapply(packages, library, character.only=TRUE)
```

## QUESTION 01: Data Visualisation for Science Communication

*Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. **Do not make a boxplot**.*

*Use the following references to guide you:*

-   [*https://www.nature.com/articles/533452a*](https://www.nature.com/articles/533452a){.uri}
-   [*https://elifesciences.org/articles/16800*](https://elifesciences.org/articles/16800){.uri}

*Note: Focus on visual elements rather than writing misleading text on it.*

### a) Provide your figure here:

```{r bad figure code, echo=FALSE, warning=FALSE}
ggplot(
  penguins,
  aes(log(flipper_length_mm), body_mass_g),
  ) + geom_point(
  color = "blue"
) +geom_smooth(
  color = "blue",
  se = FALSE,
  method = lm,
  formula = y ~ poly(x, 3)
)+ scale_y_reverse(
  labels = scales::scientific,
  n.breaks = 30
) + scale_x_continuous(
  n.breaks = 21
)
```

### b) Write about how your design choices mislead the reader about the underlying data (200-300 words).

The design of this graph is misleading for a variety of reasons, firstly the trend line is the same colour as the data points obscuring some of them. This makes it harder to read the actual points on the graph. The trend line is also over fitted to a polynomial formula implying a greater significance to the relationship than is honestly shown in the data. Moreover, the X axis is needlessly placed on a log scale compressing the data points and making the real world values harder to see. The x axis also has too many ticks and guidelines on it, many of which do not align with the data, this makes the graph much busier and seeing the values more difficult. Similarly, the y axis also contains too many ticks, and all the values are needlessly written using scientific notation reducing the ability of the graph to be understood intuitively. This can also have the effect of making the data seem more spaced out than it is. The final confusing factor is that the y axis is plotted upside down with the smallest values at the top of the graph. This has the effect of making it intuitively seem as if penguins with longer flippers weight less when in reality the reverse is true.

[Tips on How to Display Data Badly](https://sphweb.bumc.bu.edu/otlt/mph-modules/bs/datapresentation/DataPresentation9.html)

[Exploring Different Smooths](https://stats.oarc.ucla.edu/r/faq/how-can-i-explore-different-smooths-in-ggplot2/)

[Scales for Continuous Data](https://ggplot2.tidyverse.org/reference/scale_continuous.html)

------------------------------------------------------------------------

## QUESTION 2: Data Pipeline

*Write a data analysis pipeline in your .rmd RMarkdown file. You should be aiming to write a clear explanation of the steps, the figures visible, as well as clear code.*

*Your code should include the steps practiced in the lab session:*

-   *Load the data*

-   *Appropriately clean the data*

-   *Create an Exploratory Figure (**not a boxplot**)*

-   *Save the figure*

-   ***New**: Run a statistical test*

-   ***New**: Create a Results Figure*

-   *Save the figure*

*An exploratory figure shows raw data, such as the distribution of the data. A results figure demonstrates the stats method chosen, and includes the results of the stats test.*

*Between your code, communicate clearly what you are doing and why.*

*Your text should include:*

-   *Introduction*

-   *Hypothesis*

-   *Stats Method*

-   *Results*

-   *Discussion*

-   *Conclusion*

*You will be marked on the following:*

### a) Your code for readability and functionality

### b) Your figures for communication

### c) Your text communication of your analysis

*Below is a template you can use.*

------------------------------------------------------------------------

### Introduction

```{r Data Exploration}

# Make sure your code prints. 

```

### Hypothesis

### Statistical Methods

```{r Statistics}

# Make sure your code prints. 

```

### Results & Discussion

```{r Plotting Results}

# Make sure your code prints. 

```

### Conclusion

------------------------------------------------------------------------

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:*

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*
